---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(shiny)
library(bslib)

create_app_link <- function(mode = c("app", "editor"), header = TRUE) {
  mode <- match.arg(mode)

  app_code <- "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAI4uCRkAHlMZCDhqAH1OKAlBCBlCmQATKC5M6PFcQKKZMIYoCEs4CtEJQLjA9jgGADcWnLzB8T8wAF8AXSA"

  app_url <- sprintf(
    "https://shinylive.io/r/%s/#code=%s", mode, app_code
  )

  if (!header) {
    app_url <- paste0(app_url, "&h=0")
  }

  tags$iframe(
    # To allow the content to fill the full screen card
    class = "html-fill-item",
    src = app_url,
    height = "700",
    width = "100%",
    style = "border: 1px solid rgba(0,0,0,0.175); border-radius: .375rem;",
    allowfullscreen = "",
    allow = "autoplay",
    `data-external` = "1"
  )
}
```

# blockr

<!-- badges: start -->
[![ci](https://github.com/blockr-org/blockr/actions/workflows/ci.yml/badge.svg)](https://github.com/blockr-org/blockr/actions/workflows/ci.yml)
[![codecov](https://codecov.io/github/blockr-org/blockr/graph/badge.svg?token=9AO88LK8FJ)](https://codecov.io/github/blockr-org/blockr)
<!-- badges: end -->

Building blocks for data manipulation and visualization operations.

## Installation

You can install the development version of blockr from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("blockr-org/blockr")
```

## Contribute

Easiest is to run `make`, otherwise:

1. Install npm dependencies with `packer::npm_install()`
2. Build CSS by running the script in `dev/sass.R`

## Examples

```{r, eval=TRUE, echo=FALSE}
card(create_app_link("editor", FALSE), full_screen = TRUE)
```

### Create a stack

A simple stack of blocks providing a dataset selector and a filter operation:

```r
library(blockr)
library(blockr.data)

stack <- new_stack(
  data_block,
  head_block2
)

options("BLOCKR_DEV" = T)
serve_stack(stack)
```

### Modify a stack

To add a block to a stack, you can leverage `add_block`:

```r
stack <- new_stack(data_block) |>
  add_block(select_block) |>
  add_block(filter_block, 1)
```

If you specify the position (last parameter), you can add a block right 
after the given index. For instance, the above command first inserts
a select block after the data block. Then, a filter block is included
right after the data block.

Note that you can't add a block before the data block and 
before the plot block. In a later version, we may add multiple
plots per stack.

### Moving block

TBD

### Dynamically add a block

Below is an app with an interface to dynamically add blocks.
We leverage the `new_blocks` slot of the `generate_server.stack` method.
It accepts the name of the block to insert as well as its position
in the stack. 

Note that [`{blockr.demo}`](https://github.com/blockr-org/block.demo) 
utilises [`{masonry}`](https://github.com/blockr-org/masonry) to provide
a better user experience, allowing one to drag and drop blocks
within the stack.

```r
library(shiny)
library(blockr)
library(blockr.data)

blocks <- list(
  filter_block,
  select_block,
  arrange_block,
  group_by_block,
  summarize_block,
  cheat_block,
  plot_block
)

stack <- new_stack(data_block)

shinyApp(
  ui = bslib::page_fluid(
    div(
      class = "d-flex justify-content-center",
      tags$button(
        type = "button",
        "Add a new block",
        class = "btn btn-primary",
        class = "my-2",
        `data-bs-toggle` = "offcanvas",
        `data-bs-target` = "#addBlockCanvas",
        `aria-controls` = "addBlockCanvas"
      )
    ),
    generate_ui(stack, id = "mystack"),
    off_canvas(
      id = "addBlockCanvas",
      title = "My blocks",
      position = "bottom",
      radioButtons(
        "selected_block",
        "Choose a block",
        choices = c(
          "filter_block" = 1,
          "select_block" = 2,
          "arrange block" = 3,
          "group_by block" = 4,
          "summarize block" = 5,
          "cheat block" = 6,
          "plot block" = 7
        ),
        inline = TRUE
      ),
      actionButton("add", icon("plus"), `data-bs-dismiss` = "offcanvas")
    )
  ),
  server = function(input, output, session) {
    vals <- reactiveValues(new_blocks = NULL)
    o <- generate_server(
      stack,
      id = "mystack",
      new_blocks = reactive(vals$new_blocks)
    )

    observeEvent(input$add, {
      block <- blocks[[as.numeric(input$selected_block)]]
      # add_block expect the current stack, the block to add and its position
      # (NULL is fine for the position, in that case the block will
      # go at the end)
      vals$new_blocks <- list(
        block = block,
        position = length(o$stack)
      )
    })
  }
)
```

Blocks can be removed in any context, but keep in mind that 
a data block can't be removed unless it is the last stack block.

At the moment, we don't check whether removing a block breaks the pipeline.

### Example with modules

The stack can be nested within modules. This is what we do in
[`{blockr.demo}`](https://github.com/blockr-org/block.demo).

```r
library(shiny)
library(blockr.data)
library(blockr)

options(shiny.fullstacktrace = TRUE)

stack1 <- new_stack(
  data_block,
  filter_block,
  select_block
)

block_ui <- function(id) {
  ns <- shiny::NS(id)
  uiOutput(ns("block"))
}

block_server <- function(id) {
  moduleServer(
    id,
    function(input, output, session){
      ns <- session$ns

      output$block <- renderUI({
        generate_ui(stack1, id = ns("b"))
      })

      generate_server(stack1, id = "b")
    }
  )
}

ui <- function(req) {
  fluidPage(
    theme = bslib::bs_theme(5L),
    div(
      class = "row",
      block_ui("x")
    )
  )
}
server <- function(input, output, session) {
  block_server("x")
}

shinyApp(ui, server)
```
