---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(blockr)
library(bslib)
```

# blockr

<!-- badges: start -->
[![ci](https://github.com/blockr-org/blockr/actions/workflows/ci.yml/badge.svg)](https://github.com/blockr-org/blockr/actions/workflows/ci.yml)
[![codecov](https://codecov.io/github/blockr-org/blockr/graph/badge.svg?token=9AO88LK8FJ)](https://codecov.io/github/blockr-org/blockr)
<!-- badges: end -->

Building blocks for data manipulation and visualization operations.

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAI4uCRkAHlMZCDhqAH1OKAlBCBlCmQATKC5M6PFcQKKZdjhuOAJSCtEJQLjA+oYANzhshPE+HIk-MABfAF0gA",
    "editor"
  ),
  full_screen = TRUE
)
```

`{blockr}` has been built for webR (wasm) and is available for download with 
`webr::install("blockr", repos = "https://blockr-org.github.io/webr-repos")`.

```r
library(blockr)
library(blockr.data)

stack <- new_stack(
  data_block,
  select_block
)
serve_stack(stack)
```

## Installation

You can install the development version of blockr from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("blockr-org/blockr")
```

## Contribute

Easiest is to run `make`, otherwise:

1. Install npm dependencies with `packer::npm_install()`
2. Build CSS by running the script in `dev/sass.R`
