---
title: "blockr: case studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blockr: case studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(bslib)
knitr::opts_chunk$set(echo = TRUE)
```

## Why blockr?

`blockr` is an R package designed to democratize data analysis by providing a flexible, intuitive, and code-free approach to building data pipelines. It allows users to create powerful data workflows using pre-built blocks that can be easily connected, all without writing a single line of code.

### Key Features

1. **User-Friendly Interface**: Build data pipelines using a drag-and-drop interface.
2. **Flexibility**: Easily add, remove, or rearrange blocks in your pipeline.
3. **Extensibility**: Developers can create custom blocks to extend functionality.
4. **Reproducibility**: Pipelines created with `blockr` are easily shareable and reproducible, with exportable code.
5. **Interactivity**: Real-time feedback as you build and modify your pipeline.

## blockr Across Industries

The flexibility of `blockr` makes it valuable across various industries. Let's explore how it can be applied in different sectors with detailed examples. Somex examples require to create a new field such as the `new_slider_field` described in the corresponding vignette.

```{r slider_block, include=FALSE}
# Provided by John, new_range_field not working
new_slider_field <- function(value = 5L, min = 0L, max = 10L, step = 1L, ...) {
  blockr::new_field(
    value = value,
    min = min,
    max = max,
    step = step,
    ...,
    class = "slider_field"
  )
}

#' @method ui_input slider_field
#' @export
ui_input.slider_field <- function(x, id, name) {
  shiny::sliderInput(
    blockr:::input_ids(x, id),
    name,
    value = blockr::value(x, "value"),
    min = blockr::value(x, "min"),
    max = blockr::value(x, "max"),
    step = blockr::value(x, "step")
  )
}

#' @method validate_field slider_field
#' @export
validate_field.slider_field <- function(x) {
  val <- value(x)
  min <- value(x, "min")
  max <- value(x, "max")
  step <- value(x, "step")

  validate_number(val)

  if (length(min)) {
    validate_number(min, "min")
  }

  if (length(max)) {
    validate_number(max, "max")
  }

  if (length(step)) {
    validate_number(step, "step")
  }

  NextMethod()
}

#' @method ui_update slider_field
#' @export
ui_update.slider_field <- function(x, session, id, name) {
  updateSliderInput(
    session,
    blockr::input_ids(x, id),
    blockr::get_field_name(x, name),
    blockr::value(x),
    blockr::value(x, "min"),
    blockr::value(x, "max")
  )
}
```

```{r filter_block, include=FALSE}
#' Correlation plot
#' @param data dataset.
#' @param ... ignored.
#' @export
filter_in_block <- function(data, ...) {
  sub_fields <- \(data, column) {
    if (!length(column)) {
      column <- names(data)[1]
    }

    list(
      values = blockr::select_field(data[[column]][1], unique(data[[column]]), multiple = TRUE)
    )
  }

  fields <- list(
    column = blockr::new_select_field(
      \(data) names(data)[1],
      choices = \(data) names(data),
      title = "Column"
    ),
    values = blockr::new_list_field(character(), sub_fields)
  )

  blockr::new_block(
    fields = fields,
    expr = quote({
      vals <- unname(.(values))
      data |>
        dplyr::filter(!!dplyr::sym(.(column)) %in% vals[[1]])
    }),
    ...,
    class = c("filter_in_block", "transform_block", "submit_block")
  )
}

blockr::register_block(
  constructor = filter_in_block,
  name = "Filter in",
  description = "Filter on a vector",
  classes = c("filter_in_block", "transform_block"),
  input = "data.frame",
  output = "data.frame"
)
```


### 1. Finance: Stock Price Forecasting

In this example, we'll create a pipeline that fetches recent stock data using the [quantmod package](https://github.com/joshuaulrich/quantmod), performs time series analysis, and forecasts future stock prices using the [Prophet model](https://github.com/facebook/prophet). We first design the `stock_data_block`, containing a field to select the stock items and generate the data. The `prophet_forecast_block` does all the modeling part.

```{r stock_example}
library(blockr)
library(quantmod)
library(prophet)

# Custom block to fetch stock data
stock_data_block <- function(...) {
  # stocks to pick (top 10)
  pick_stock <- \()  c("NVDA", "TSLA", "AAPL", "MSFT", "AVGO", "AMZN", "AMD", "PLTR", "TSM", "META")


  new_block(
    fields = list(
      ticker = new_select_field(pick_stock()[1], pick_stock, multiple = FALSE, title = "Ticker")
    ),
    expr = quote({
      data_xts <- getSymbols(.(ticker), src = "yahoo", auto.assign = FALSE)
      data.frame(Date = index(data_xts), coredata(data_xts)) |>
        tail(700) # only considering last 700 days for this example
    }),
    class = c("stock_data_block", "data_block"),
    ...
  )
}

# Custom block for Prophet forecasting
prophet_forecast_block <- function(columns = character(), ...) {
  all_cols <- function(data) colnames(data)[2:length(colnames(data))]


  new_block(
    fields = list(
      # date_col = new_select_field(columns, all_cols, multiple=FALSE, title="Date"),
      value_col = new_select_field(columns, all_cols, multiple = FALSE, title = "Value"),
      periods = new_slider_field(7, min = 0, max = 365, title = "Forecast duration")
    ),
    expr = quote({
      df <- data.frame(
        ds = data$Date,
        y = data[[.(value_col)]]
      )
      model <- prophet(df)
      future <- make_future_dataframe(model, periods = .(periods))
      forecast <- predict(model, future)
      plot(model, forecast)
    }),
    class = c("prophet_forecast_block", "plot_block"),
    ...
  )
}

# Register custom blocks
register_block(
  stock_data_block,
  name = "Stock Data",
  description = "Fetch stock data",
  classes = c("stock_data_block", "data_block"),
  input = NA_character_,
  output = "data.frame"
)
register_block(
  prophet_forecast_block,
  name = "Prophet Forecast",
  description = "Forecast using Prophet",
  classes = c("prophet_forecast_block", "plot_block"),
  input = "data.frame",
  output = "plot"
)

# Create the stack
stock_forecast_stack <- new_stack(
  stock_data_block(),
  prophet_forecast_block()
)

serve_stack(stock_forecast_stack)
```
 

### 2. Pharmaceutical: Clinical Trial Analysis

#### 2.1 AE Forest Plot

This forest plot visualizes the relative risk of adverse events between two treatment arms in a clinical trial. In this case, it compares "Xanomeline High Dose" to "Xanomeline Low Dose" starting from the [pharmaverseadam](https://github.com/pharmaverse/pharmaverseadam) `adae` dataset.

```{r ae_forest}
library(dplyr)
library(tidyr)
library(forestplot)

# Function to create adverse event forest plot
create_ae_forest_plot <- function(data, usubjid_col, arm_col, aedecod_col, n_events) {
  data <- data |> filter(.data[[arm_col]] != "Placebo")
  # Convert column names to strings
  usubjid_col <- as.character(substitute(usubjid_col))
  arm_col <- as.character(substitute(arm_col))
  aedecod_col <- as.character(substitute(aedecod_col))

  # Calculate the total number of subjects in each arm
  n_subjects <- data |>
    select(all_of(c(usubjid_col, arm_col))) |>
    distinct() |>
    group_by(across(all_of(arm_col))) |>
    summarise(n = n(), .groups = "drop")

  # Calculate AE frequencies and proportions
  ae_summary <- data |>
    group_by(across(all_of(c(arm_col, aedecod_col)))) |>
    summarise(n_events = n_distinct(.data[[usubjid_col]]), .groups = "drop") |>
    left_join(n_subjects, by = arm_col) |>
    mutate(proportion = n_events / n)

  # Select top N most frequent AEs across all arms
  top_aes <- ae_summary |>
    group_by(across(all_of(aedecod_col))) |>
    summarise(total_events = sum(n_events), .groups = "drop") |>
    top_n(n_events, total_events) |>
    pull(all_of(aedecod_col))

  # Get unique treatment arms
  arms <- unique(data[[arm_col]])
  if (length(arms) != 2) {
    stop("This plot requires exactly two treatment arms.")
  }
  active_arm <- arms[1]
  control_arm <- arms[2]

  # Filter for top AEs and calculate relative risk
  ae_rr <- ae_summary |>
    filter(.data[[aedecod_col]] %in% top_aes) |>
    pivot_wider(
      id_cols = all_of(aedecod_col),
      names_from = all_of(arm_col),
      values_from = c(n_events, n, proportion)
    ) |>
    mutate(
      RR = .data[[paste0("proportion_", active_arm)]] / .data[[paste0("proportion_", control_arm)]],
      lower_ci = exp(log(RR) - 1.96 * sqrt(
        1 / .data[[paste0("n_events_", active_arm)]] +
          1 / .data[[paste0("n_events_", control_arm)]] -
          1 / .data[[paste0("n_", active_arm)]] -
          1 / .data[[paste0("n_", control_arm)]]
      )),
      upper_ci = exp(log(RR) + 1.96 * sqrt(
        1 / .data[[paste0("n_events_", active_arm)]] +
          1 / .data[[paste0("n_events_", control_arm)]] -
          1 / .data[[paste0("n_", active_arm)]] -
          1 / .data[[paste0("n_", control_arm)]]
      ))
    )

  # Prepare data for forest plot
  forest_data <- ae_rr |>
    mutate(
      label = paste0(
        .data[[aedecod_col]], " (",
        .data[[paste0("n_events_", active_arm)]], "/", .data[[paste0("n_", active_arm)]], " vs ",
        .data[[paste0("n_events_", control_arm)]], "/", .data[[paste0("n_", control_arm)]], ")"
      )
    )

  # Create forest plot
  forestplot::forestplot(
    labeltext = cbind(
      forest_data$label,
      sprintf("%.2f (%.2f-%.2f)", forest_data$RR, forest_data$lower_ci, forest_data$upper_ci)
    ),
    mean = forest_data$RR,
    lower = forest_data$lower_ci,
    upper = forest_data$upper_ci,
    align = c("l", "r"),
    graphwidth = unit(60, "mm"),
    cex = 0.9,
    lineheight = unit(8, "mm"),
    boxsize = 0.35,
    col = fpColors(box = "royalblue", line = "darkblue", summary = "royalblue"),
    txt_gp = fpTxtGp(label = gpar(cex = 0.9), ticks = gpar(cex = 0.9), xlab = gpar(cex = 0.9)),
    xlab = paste("Relative Risk (", active_arm, " / ", control_arm, ")"),
    zero = 1,
    lwd.zero = 2,
    lwd.ci = 2,
    xticks = c(0.5, 1, 2, 4),
    grid = TRUE,
    title = paste("Relative Risk of Adverse Events (", active_arm, " vs ", control_arm, ")")
  )
}

forest_plot_block <- function(...) {
  new_block(
    fields = list(
      usubjid_col = new_select_field(
        "USUBJID",
        function(data) colnames(data),
        multiple = FALSE,
        title = "Subject ID Column"
      ),
      arm_col = new_select_field(
        "ACTARM",
        function(data) colnames(data),
        multiple = FALSE,
        title = "Treatment Arm Column"
      ),
      aedecod_col = new_select_field(
        "AEDECOD",
        function(data) colnames(data),
        multiple = FALSE,
        title = "AE Term Column"
      ),
      n_events = new_numeric_field(
        10,
        min = 5, max = 20, step = 1,
        title = "Number of Top AEs to Display"
      )
    ),
    expr = quote({
      create_ae_forest_plot(data, .(usubjid_col), .(arm_col), .(aedecod_col), .(n_events))
    }),
    class = c("adverse_event_plot_block", "plot_block"),
    ...
  )
}

# Register the custom block
register_block(
  forest_plot_block,
  name = "Adverse Event Forest Plot",
  description = "Create a forest plot of adverse events comparing two treatment arms",
  classes = c("adverse_event_plot_block", "plot_block"),
  input = "data.frame",
  output = "plot"
)

# Create the stack
clinical_trial_stack <- new_stack(
  new_dataset_block(
    selected = "adae",
    package = "pharmaverseadam"
  ),
  # filter_in_block(),
  forest_plot_block()
)

serve_stack(clinical_trial_stack)
```
#### 2.2 Demographics Table

This demographics table is taken from the `cardinal` package of [FDA Safety Tables and Figures](https://github.com/pharmaverse/cardinal) and demonstrates `gt` and `rtables` outputs starting from the [pharmaverseadam](https://github.com/pharmaverse/pharmaverseadam) `adsl` dataset.


```{r cardinal_tables}
library(shiny)
library(blockr)
library(blockr.falcon)
library(blockr.pharmaverseadam)

cardinal02_block <- function(...) {
  all_cols <- function(data) colnames(data)

  fields <- list(
    columns = new_select_field(
      c("SEX", "AGE", "AGEGR1", "RACE", "ETHNIC", "COUNTRY"),
      all_cols,
      multiple = TRUE,
      title = "Variables"
    )
  )

  expr <- quote({
    data <- droplevels(data)

    rtables <- cardinal::make_table_02(
      df = data,
      vars = .(columns)
    )

    gt <- cardinal::make_table_02_gtsum(
      df = data,
      vars = .(columns)
    )

    list(
      rtables = rtables,
      gt = gt
    )
  })

  blockr::new_block(
    expr = expr,
    fields = fields,
    ...,
    class = c("cardinal02_block", "rtables_block", "submit_block")
  )
}

blockr::register_block(
  cardinal02_block,
  "Cardinal 02",
  "A Cardinal 02 table",
  input = "data.frame",
  output = "list",
  classes = c("cardinal02_block", "rtables_block", "submit_block")
)


# Create the stack
rtables_stack <- new_stack(
  new_adam_block(selected = "adsl"),
  cardinal02_block()
)

serve_stack(rtables_stack)
```


### 3. Environmental Science: Air Quality Analysis and Prediction

This example demonstrates a pipeline for analyzing air quality data and predicting future pollution levels using actual data from the [openair](https://github.com/cran/openair) package.
This pipeline imports actual air quality data from the `openair` package and forecasts future pollution levels using an ARIMA model.

```{r environment}
library(blockr)
library(openair)
library(forecast)


# Custom block for air quality data import
air_quality_block <- function(...) {
  new_block(
    fields = list(
      site = new_select_field(
        "kc1",
        \() openair::importMeta()$code,
        multiple = FALSE,
        title = "Monitoring Site"
      ),
      start_year = new_numeric_field(
        2020,
        min = 1990,
        max = as.numeric(format(Sys.Date(), "%Y")),
        step = 1,
        title = "Start Year"
      ),
      end_year = new_numeric_field(
        as.numeric(format(Sys.Date(), "%Y")),
        min = 1990,
        max = as.numeric(format(Sys.Date(), "%Y")),
        step = 1,
        title = "End Year"
      )
    ),
    expr = quote({
      importAURN(site = .(site), year = .(start_year):.(end_year)) |> tail(700)
    }),
    class = c("air_quality_block", "data_block"),
    ...
  )
}


# Custom block for pollution forecasting
pollution_forecast_block <- function(columns = character(), ...) {
  all_cols <- function(data) setdiff(colnames(data), c("date", "site", "source"))

  new_block(
    fields = list(
      pollutant = new_select_field(columns, all_cols, multiple = FALSE, title = "Pollutant"),
      horizon = new_slider_field(
        30,
        min = 1,
        max = 365,
        step = 1,
        title = "Forecast Horizon (days)"
      )
    ),
    expr = quote({
      ts_data <- ts(na.omit(data[[.(pollutant)]]), frequency = 365)
      model <- auto.arima(ts_data)
      forecast_result <- forecast(model, h = .(horizon))
      plot(forecast_result, main = paste("Forecast of", .(pollutant), "levels"))
    }),
    class = c("pollution_forecast_block", "plot_block"),
    ...
  )
}

# Register custom blocks
register_block(
  air_quality_block,
  name = "Air Quality Data",
  description = "Import air quality data",
  classes = c("air_quality_block", "data_block"),
  input = NA_character_,
  output = "data.frame"
)

register_block(
  pollution_forecast_block,
  name = "Pollution Forecast",
  description = "Forecast pollution levels",
  classes = c("pollution_forecast_block", "plot_block"),
  input = "data.frame",
  output = "plot"
)

# Create the stack
air_quality_stack <- new_stack(
  air_quality_block(),
  pollution_forecast_block(columns = "no2")
)

serve_stack(air_quality_stack)
```
### 4. Marketing: Causal Impact Analysis of Marketing Interventions

This example demonstrates how to use [CausalImpact](https://google.github.io/CausalImpact/CausalImpact.html) to analyze the effect of marketing interventions on sales data. 
This pipeline generates dummy marketing data with an intervention, then uses CausalImpact to analyze the effect of the intervention on sales.

```{r date_field, include=FALSE}
library(shiny)
new_date_field <- function(value = Sys.Date(), min = NULL, max = NULL, ...) {
  blockr::new_field(
    value = value,
    min = min,
    max = max,
    ...,
    class = "date_field"
  )
}

date_field <- function(...) {
  validate_field(new_date_field(...))
}

#' @method ui_input date_field
#' @export
ui_input.date_field <- function(x, id, name) {
  ns <- NS(input_ids(x, id))

  shiny::dateInput(
    blockr:::input_ids(x, id),
    name,
    value = blockr::value(x, "value"),
    min = blockr::value(x, "min"),
    max = blockr::value(x, "max")
  )
}

#' @method validate_field date_field
#' @export
validate_field.date_field <- function(x, ...) {
  x
}

#' @method ui_update date_field
#' @export
ui_update.date_field <- function(x, session, id, name) {
  updateDateInput(
    session,
    blockr:::input_ids(x, id),
    blockr:::get_field_name(x, name),
    value = blockr::value(x),
    min = blockr::value(x, "min"),
    max = blockr::value(x, "max")
  )
}
```




```{r intervention_impact}
library(blockr)
library(CausalImpact)
library(dplyr)

# Custom block to load and prepare marketing data
new_marketing_data_block <- function(...) {
  new_block(
    fields = list(
      start_date = date_field(
        Sys.Date() - 365,
        min = Sys.Date() - 730,
        max = Sys.Date() - 1,
        label = "Start Date"
      ),
      intervention_date = date_field(
        Sys.Date() - 180,
        min = Sys.Date() - 729,
        max = Sys.Date(),
        label = "Intervention Date"
      ),
      end_date = date_field(
        Sys.Date(),
        min = Sys.Date() - 364,
        max = Sys.Date(),
        label = "End Date"
      )
    ),
    expr = quote({
      # Generate dummy data for demonstration
      dates <- seq(as.Date(.(start_date)), as.Date(.(end_date)), by = "day")
      sales <- cumsum(rnorm(length(dates), mean = 100, sd = 10))
      ad_spend <- cumsum(rnorm(length(dates), mean = 50, sd = 5))

      # Add intervention effect
      intervention_index <- which(dates == as.Date(.(intervention_date)))
      sales[intervention_index:length(sales)] <- sales[intervention_index:length(sales)] * 1.2

      data.frame(
        date = dates,
        sales = sales,
        ad_spend = ad_spend
      )
    }),
    class = c("marketing_data_block", "data_block"),
    ...
  )
}

# Custom block for CausalImpact analysis
new_causal_impact_block <- function(columns = character(), ...) {
  all_cols <- function(data) colnames(data)[2:length(colnames(data))]

  new_block(
    fields = list(
      response_var = new_select_field(
        columns,
        all_cols,
        multiple = FALSE,
        title = "Response Variable"
      ),
      covariate_var = new_select_field(
        columns,
        all_cols,
        multiple = FALSE,
        title = "Covariate Variable"
      ),
      pre_period_end = date_field(
        Sys.Date() - 181,
        min = Sys.Date() - 729,
        max = Sys.Date() - 1,
        label = "Pre-Period End Date"
      ),
      post_period_start = date_field(
        Sys.Date() - 180,
        min = Sys.Date() - 728,
        max = Sys.Date(),
        label = "Post-Period Start Date"
      )
    ),
    expr = quote({
      data <- data.frame(
        y = data[[.(response_var)]],
        x = data[[.(covariate_var)]]
      )

      pre_period <- c(min(data$date), as.Date(.(pre_period_end)))
      post_period <- c(as.Date(.(post_period_start)), max(data$date))

      impact <- CausalImpact(data, pre_period, post_period)
      plot(impact)
    }),
    class = c("causal_impact_block", "plot_block"),
    ...
  )
}

# Register custom blocks
register_block(
  new_marketing_data_block,
  name = "Marketing Data",
  description = "Load and prepare marketing data",
  classes = c("marketing_data_block", "data_block"),
  input = NA_character_,
  output = "data.frame"
)
register_block(
  new_causal_impact_block,
  name = "Causal Impact Analysis",
  description = "Perform Causal Impact analysis on marketing data",
  classes = c("causal_impact_block", "plot_block"),
  input = "data.frame",
  output = "plot"
)

# Create the stack
marketing_impact_stack <- new_stack(
  new_marketing_data_block(),
  new_causal_impact_block()
)

serve_stack(marketing_impact_stack)
```


### 5. Dynamical systems

In the below example, we implemented the Lorenz attractor and solve it with the `{pracma}` R package (technically, the reason using `{pracma}` over `{deSolve}` or `{diffeqr}` is because only `{pracma}` is available for shinylive required by the embeded demo).

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZBQWAAWpUqnaIA9IZFjJAWiIMA5hmstSWgK50MLIodqNz8xe2kyQjp6Bsa+RBgM5tRQ7DAYVtZCAJTJAhBeTKwcXLyaphIMdtaooqQATAHhymoaQbr6RiaihZY2dg7Oru6e9FHVAfUhTeGR0bHxiSlpEOncLIxQDACefAWSM-OLK3yoDFAEMFCbC-s760XWJWXlM+kQcNQA+kQAJnBP6zIAPOYyAGZOCAEUjuCB8DCQ5IyEDpdIyGSieQQABePz+gOBoJIMj4pFky1kqCWMHY0NhEARCLqlKpCIAGqoZMSGKTgABGAC6MgAVDJlhzuQBqfnAcrcvkCgDMnNw8LpMgAmkyWWzxbzcQL1X9pZzknLaXSAFpM8wCrkarXC5kk9jAGWWsXcnX2znyhEzBEAXzhtP+LDg3Feyl+iJYnEEhpkUCZD2eECc8AYLAIT39gdefHMAA5DFLZOZ2QAGWTlIv690yOixx5PBNJlNpgNBrPFgsAVhLMmL5YNdIINfjibgydT6Zb5WzsnZ06L5fld1pcc+LXEkbp4+DTM37D7VLgAA89kz2C5OA4nOR1wrYhhXlAuBh-vt4NeFTI3nAACztt-vgFArIlbvssRZMhojJqDOSpMtBJpQRWUbvqQYFqCWwEKqQ-rQNwsFznu-42qyNTqBCfDHLIGBrPqMhUQQqQYR6GGIe+8wRpiTJIhQKKelSLFUpCGAEeo3CxCRdRgJ+K5mIMYD3lw0kSCkC7pD6sw5Ac4jojIy6cJp17LlJ6xAUutZXKURCkIpa6VpiA61JoB6ycsKQmXS7z-FATjcFZxDcImHDgZoywYOyzkYJUYC8TR8rLtYcBEDAihsFZ6zpDM7AjgAbh8ekSHweXiMkYBepyQA", # nolint
    "app",
    mode = "editor"
  ),
  full_screen = TRUE
)
```

<p class="text-center">
<a 
class="btn btn-primary" 
data-bs-toggle="collapse" 
href="#lorenz-mod" 
role="button" 
aria-expanded="false" 
aria-controls="lorenz-mod">
Toggle code
</a>
</p>

<div class="collapse" id="lorenz-mod">

```r
library(blockr)
library(pracma)
library(blockr.ggplot2)

new_ode_block <- function(...) {

  lorenz <- function (t, y, parms) {
    c(
      X = parms[1] * y[1] + y[2] * y[3],
      Y = parms[2] * (y[2] - y[3]),
      Z = -y[1] * y[2] + parms[3] * y[2] - y[3]
    )
  }

  fields <- list(
    a = new_numeric_field(-8/3, -10, 20),
    b = new_numeric_field(-10, -50, 100),
    c = new_numeric_field(28, 1, 100)
  )

  new_block(
    fields = fields,
    expr = substitute(
      as.data.frame(
        ode45(
          fun, 
          y0 = c(X = 1, Y = 1, Z = 1),
          t0 = 0,
          tfinal = 100,
          parms = c(.(a), .(b), .(c))
        )
      ),
      list(fun = lorenz)
    ),
    ...,
    class = c("ode_block", "data_block")
  )
}

stack <- new_stack(
  new_ode_block,
  # Coming from blockr.ggplot2
  new_ggplot_block(
    func = c("x", "y"), 
    default_columns = c("y.1", "y.2")
  ), 
  new_geompoint_block
)
serve_stack(stack)
```
</div>
