---
title: "3. Case study: create a plot block"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Case study: create a plot block}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)

note_box <- function(..., color) {
  htmltools::div(
    class = sprintf(
      "m-3 p-2 text-%s-emphasis bg-%s-subtle border border-%s-subtle rounded-3",
      color, color, color
    ),
    ...
  )
}
```

In this vignette, we show how to create a plot block layer by layer with the `{ggplot2}` package.

```{r, eval=TRUE, echo=FALSE}
note_box(
  "Before going further, please be sure to read the S3 chapter of advanced R: https://adv-r.hadley.nz/s3.html.", #nolint
  color = "warning"
)
```

## Introduction

### Anatomy of a `{ggplot2}` plot

With `{ggplot2}`, plots are built __layer__ by __layer__. It all starts with
`ggplot()` which initialises the ggplot object passing optional data and mappings.
Then, we add __geoms__, like `geom_point()`. Geoms also accept custom mappings (in case not mapping is 
passed through the first `ggplot()` call), data as well as other parameters.
At the end of the day, there are many ways to build a ggplot:

```r
# 1
ggplot(data = <DATA>, mapping = <MAPPING>) +
  geom_point()
# 2
ggplot(data = <DATA>) +
  geom_point(mapping = <MAPPING>) +
  geom_point(mapping = <MAPPING>)

# 3
ggplot() +
  geom_point(data = <DATA>, mapping = <MAPPING>) +
  geom_point(data = <DATA>, mapping = <MAPPING>)
```

That said, even if we pass data and mappings to `ggplot()`, we can still overwrite them in
subsequent calls to `geom_*()` functions. Therefore, we'll go for the first option.

### To do list

What do we need to create?

It seems obvious to add a new `ggplot_block()` __constructor__. On the Shiny side, we have to handle the 
plot output element. Remember that the `server_output()` __generic__ is defined in the `{blockr}`
`server.R` script. It supports tables with `server_output.block()` and plots with `server_output.plot_block()`.
Therefore, we don't need to create another S3 method and have to make sure that our new plot block inherits
from the `plot_block` class to __dispatch__ to the correct method. Most of our work will be on the block side since
we'll have to generate the plot layer by layer, where each layer is a block.

```{r, eval=TRUE, echo=FALSE}
note_box(
  "Notice that the current {blockr} plot_block isn't built layer by layer but as a huge monolitic structure", #nolint
  "which is certainly less flexible.",
  color = "primary"
)
```

### Create the ggplot block

To create a new block, we call the `new_block()` constructor. It expects:

- __fields__: a list of field, which are translated into shiny inputs.
- __expr__: the expression return by the block, necessary to produce an output and export the code.
- __name__: a name (randomely choosen), you don't need to worry about this).
- __class__: a class to dispatch to the relevant S3 methods.
- __layout__: an optional layout (default to `default_layout_fields()`).

Our block will have `c("ggplot_block", "plot_block")` as classes, some fields to pass mappings and we
leave the layout to the default choice. Note that the mapping field depends on the provided data. 
To stay simple, we assume to only handle `x` and `y` aesthetics.

```{r}
new_ggplot_block <- function(data, ...) {
  new_block(
    fields = list(
      # TO DO
    ),
    expr = quote(),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}

ggplot_block <- function(data, ...) {
  initialize_block(new_ggplot_block(data, ...), data)
}
```

We finally end with `new_ggplot_block` (internal constructor) and 
`ggplot_block` (user facing constructor). The next step is to create the mappings fields with 
2 `select_field()`. We define the `create_mapping_field` helper:

```r
create_mapping_field <- function(data) {
  cols <- colnames(data)
  new_select_field(value = cols[1], cols)
}
```

We modify `new_ggplot_block` such that:

```r
new_ggplot_block <- function(data, ...) {

  new_block(
    fields = list(
      x = create_mapping_field(data),
      y = create_mapping_field(data)
    ),
    expr = quote(),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}
```

It is now time to produce the expression, which is certainly the most technical part as it involves a bit of metaprogramming. 
The expression must not be evaluated in the block, that's why it is wrapped in a `quote()`. 
`{blockr}` is then able to generate the expression with `generate_code()` and evaluate it with `evaluate_block()`.


We start by defining a naive implementation of the `ggplot_expr` function. If we were directly doing this:

```{r}
ggplot_expr <- function(data, x, y) {
  quote(ggplot(mapping = aes(x = .data[[.(x)]], y = .data[[.(y)]])))
}
```

we would not be able to capture the user input `x` and `y` and therefore, the generated expression wouldn't be valid as
we want to substitute by the input values of `x` and `y`. This is why we leverage `bquote()`: it only evaluates arguments wrapped
by `.()` with variables in the environment, which makes it easier to use than `substitute()` (since we explicitly mark what we want to
evaluate).

```{r}
ggplot_expr <- function(data, x, y) {
  bquote(
    ggplot(mapping = aes(x = .data[[.(x)]], y = .data[[.(y)]])),
    list(x = x, y = y)
  )
}
```

We need to create a specific field which contains the expression with `new_hidden_field`, which won't be visible
to the user. This eventually gives us:

```r
new_ggplot_block <- function(data, ...) {
  new_block(
    fields = list(
      x = create_mapping_field(data),
      y = create_mapping_field(data),
      expression = new_hidden_field(ggplot_expr)
    ),
    expr = quote(.(expression)),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}
```

You might wonder why we are doing `quote(.(expression))`. This is because `generate_code` calls `bquote()`.

### Create a geom block

Now that we have a valid `ggplot_block`, we want to add it a geometry. The easiest one is `geom_point()`.
We initialise the geom block:

```r
new_geompoint_block <- function(plot, mappings, ...) {

  geompoint_expr <- function(color) {
    bquote(geom_point(color = .(color)), list(color = color))
  }

  new_block(
    fields = list(
      color = new_select_field("blue", c("blue", "green", "red")),
      expression = new_hidden_field(geompoint_expr)
    ),
    expr = quote(.(expression)),
    class = c("plot_layer_block", "plot_block"),
    ...
  )
}

geompoint_block <- function(data, ...) {
  initialize_block(new_geompoint_block(data, ...), data)
}
```

Note the class `plot_layer_block`. This is necessary to invoke the right `evaluate_block` 
method (to use `+` instead of `%>%`).
