---
title: "1. Data blocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Data blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)
library(bslib)

note_box <- function(..., color) {
  htmltools::div(
    class = sprintf(
      "m-3 p-2 text-%s-emphasis bg-%s-subtle border border-%s-subtle rounded-3",
      color, color, color
    ),
    ...
  )
}
```

## Introduction

Data blocks are the starting point of any __stack__ and are required by any __transform__ or __plot__ block.

### Package based data block

`{blockr}` introduces a `dataset_block()` which takes as input any package containing a (or multiple) dataset(s).
This block is constructed as follows:

```r
new_dataset_block <- function(
  ...,
  dat = as.environment("package:datasets"),
  selected = character()
) {
  # code  
}
```

By default, the selected dataset is from the `datasets` package but you can actually provide any dataset you like by changing the `dat` parameter
(in the following we leverage the `{blockr.data}` datasets with the lab data pre-selected):

```r
# Define the custom data block
custom_data_block <- function(...) {
  dataset_block(
    dat = as.environment("package:blockr.data"),
    selected = "lab",
    ...
  )
}

stack <- new_stack(
  custom_data_block, 
  select_block
)
serve_stack(stack)
```

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAKCQhjDI6MkMABMoLjjAggdOIhgAfQyuQpSZAB5TGQAzBwgCUlcIPgw2vxkQQJkZEqgy0QlBCB6ektUZKHYMCgA3FgYSeDJBMFQoCShLOEQUhnTMqH9cbtH2OG44Rrg0iaFuKDppU562jFO4gF9AwM4N8Uq1QgcGohT+Q1OuXyRT6AxMslO50ujThEkCcXODFmcDBXCG4PEfjAnwAukA", #nolint
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```

```{r, eval=TRUE, echo=FALSE}
note_box(
  "In the registry vignette, we see how to register a block in a cleaner way.",
  color = "primary"
)
```

In practice, you'll always call `dataset_block()` and not `new_dataset_block()` (internal constructor).

Overall, this block is good for testing or if you get your own data package. In the next part, we see how to import
data from a file.

### Reading data from files

`{blockr}` exposes 2 extra data blocks which have to be combined with one 
of the 5 following data __parser__ blocks:

- `xpt_block`, reads xpt files.
- `sas_block`, reads sas7bdat files.
- `rds_block`, reads rds files.
- `json_block`, reads json files.
- `csv_block`, reads csv files.

#### Upload data

If you want to load data from any location on your computer, `upload_block()` is what you need.
Since the `upload_block()` temporarily moves the data in a custom location, we recognize that it might not be always possible
for security reasons.

```r
library(blockr)

stack <- new_stack(
  upload_block,
  csv_block,
  select_block
)
serve_stack(stack)
```

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAIhAzigJGQAeUxkIOGoAfUSJQQgZGQdUUSgAE2zo8VxAooJ2ADcq0Qlawpl2OG44AlIWk0C4roZGuFyufLzxOLAAXwBdIA", #nolint
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```

#### Files browser block

`filesbrowser_block()` can read data from the __server file system__.
It means that if no data is available on the server, you won't see anything.
You can customize the accessible locations as follows, through the `volumes` parameter:

```r
custom_filesbrowser_block <- function(...) {
  filesbrowser_block(
    volumes = c(vol1 = "<PATH1>", vol2 = "<PATH2>"),
    ...
  )
}
```

In the following demonstration, we uploaded a dummy dataset on the webR file system so you can play with
the block. That said, you can definitely upload your own data thanks to the corresponding
shinylive feature (upload button).

```r
library(blockr)

stack <- new_stack(
  filesbrowser_block,
  csv_block,
  select_block
)
serve_stack(stack)
```

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAIhAzigJGQAeUxkIOGoAfUSJQQgZGQAzFm44dkYianY4Bmzo8VxAooJ2ADcG0Qlmwpla8oJSLpNAuNqGdrhcrny88TiwAF9ccGh4KgB5bnCYVBY27IBxWAqMNva8QhJySmQ4ABM7VwhcR+eSbJYHt6fSF+yDgY3Fw4TgYVwBDs4UhRAcZAi2WK3CglkBwNh8NIiIgYlwiQYwweUHIbwgD2yxNJxD2cH+-0+VLguBY7AAEnBuA9AgBGAAcAE4AGwyADKDhg8AYMm2u32bVwPNw+neDI4+iV-OFuAAgvY2FBlSwYKiKvpkaj2PojlgAKIYVAQSy4G222QinWoBhlRUAVhknu9IJFgGQCGQ8gDMAa9Pt5AoADPGxRKpTKdnsDuxcAAmZWql5W3M8hPx3AABTCrKNJssZotlitADEsDqHU7cM2dbhuzyACwyACyUHCMjD2b5MqGRDodV6xfj-fFkrqablmdwEbzfwL+k3897uFFpAwMgAMnCq-pjaarfWrQBVUWtx3Ox9dvuD4fd8MyABSDh2UcZGzEUADkiCmGAZwYOcEwnJdU1lDMFV9LcPnVVD5z5XBz3JEhqxvc0UQba0ACEsDbZ0jnIxUo3-HZu2zAB2aMgyAiMeUnUhp1nOMeWzZNl2lJD5SzIU0LVK1xOLfjD24iQNCIbgYAI2tb2Iq1RQAdXtF9Dx090-wAmFmKMhiZH9IcRzHFj6JhPiRQQlcRPXX1SxVbcSCtNzFQFHlxNIupgleK8azrDTrVtCi9KOKLu1wQAUAlwcDSBkDROQeGQHgcOAZG4mRqDCONsyTJzhPTUTcCYiSd2q4sSt1MhaAYVAKFU8LLX0UjbVPSjcG609FX7HUHEsBxOA3JMRrGibsyjQMyiA-ixQUcgoN4iB6sXFNnIq9dsI89CrWwrby0rLNQsIu99E7PrO1kf07JzWzjNkfsrKA0ynuK+CdvKtcFQFGqvP0IH6uwnUYE4OpiRUy61KIzrQNtAARPrkZRnMJye5bpvG0hFRYj6w1x0b8dgiMBLK1dkKzHl3PzEH6d8ymcKUAMnU5Cp2vUzq3z6t9JrMmEPzxiaIyTOyluGsmJrjCNHL+mnKp5JVDskjVNQFBX+qCtgecRkjYui9tjdkTixYJ-yA1lq2lo9W3YN7UqlZclC1cZjCted3AOW4dg2HEFgDeuxsAElQLuiP4qSlK0oyrKcryogCqKzaBV7bahOV1zc3VndfSLDOD1wh58PhjqjfIvrqKwGPkqIVL0q5RPcvywqpHT3tfuzt26bzz2rVk4tu7ZvCQuvBHrtrmuaOzAVhaGm2ZoJkDhel5fybjQvBMQvaFUjYH1UP4tC99zkA4gIOQ4i8PI70u-fMXmR5sdj8pbDF-N7l9PfUV3v950wPPnJmB5T7iQHJyOgcIGCZBvp1HUj4+qINFDmAS4FILQVkBOFGcACBwHWjBbM6CIIEOgkBGQOC8FkI2sWIULsAEAzpqhEBx9MLClLFgIgJp4EkTDgAFWfO2ARXZsz+ktoqTiopVo0KIeI22S0pEyMIbBIUWc95MMVOJVhQ9pLCgPPwog4hwhEF4VaX8ZZ77tgsaBRUSZNhTiwdmfsDjuJYNVp9FxjjaHCh7ho2mhMj5Dzqr43AABhFgDhiSZQeLlAcABLgAHgcUxFdeYkQHLaAAGn1TJWTFQCVcTxIhLEinuIEjZLixTYJMSpq7QBioDqDw1CdAUtTcADgAD8QAIOlCeYV0lNiwEcO6wycwOxXpIlaqA1pOImfjRR0zZk+KYv-fxKsgY6I1GDNpECbgMHiTwMx+gwk6isc6U5tieQvXMhbN+2NjIULuZMuMfIGHrPXA1LZDVixvI6UocQ7RDRpMNg+LAooBbgqfk9L+EiSrr0-q-F56c+TqN2po7MHtPLqkxb5VFbNlA6k5uUC6k9K4PifALJ8WNF6k0mTIEUH9wwCQka8vx6KAnEKCfoLlvzsLSLhCCEF10ADSmwTbOjFXXa5SzZGyEKd4mCMrpEzNkRQhVbifECjqYwzlm5vl7gFNq-qYR8GiGgMc20ooyx9StWWHMj1XoyAXnC-sTLnXfwJnGYUu8OWVWcdygNxZtR6hRGQYFZLBn6H5npQWxZn4y0mfC91iat4QBKomX1-1OUsOaWInMiZSyinCA8TIMJhURRQcgx8fpZUqJ-GU2ckY63kM-pxRtMFAgZrRdm-12i83iW7bqfUFqK2dVdDXO0tFPU5hdW-TiY453Iozey3trl9XNN9JuFdusWAACs2DOjHSRMJbJznhLPdgmdziZ2hmAqmn+JVlrUz7rgX0wDN0HifbmUu5dI2grIhKl0c8bkixZY7MRCLmUzq7fGa2L6Gm+k2ZuoGT7xJYFcFlXKv4oCZBYAwVJ-7rrkSEc6EjD1r3PPJlGMcVHH3xnhQhzRQosVHX0CxgtDVDHGMIwMgDNi+o2JzHRR5JVcWUM9cBBjSpMVQYk5bKTuLYM3qYwEoUG7sVSW3Qxg8FZvSkr46HFsoyuzdiSsp1dOcFRqO5TZsT2FzyEuJdzY9FLSO4EFmZ3oJVKZZqs1mCM7StmBdzD53MpF9N0Fw8yVz+gq16RQfXYq-YtJsHIGuhU89A07IDWEjQsASAsEScc26D8Ww5jdbhhwYQYQWRkI2egDBqsRHGX+KrNWKH9ga4wZr9l07jhkKlsgfr1wS25WN8Gh5cPiFSgOCwdgABexzRT3jDn1FbYcpndaazVwm9XGu9amSTBe23DvywEkN9L-mNysY1hxFmP6oDiFymWFEUIfixZje2QWMgusHd28207u3+wk39EDlr8sRSXZGwqVmwWi46xOAwGArJemmD07cIODBeltVi8bGucUGX7Z6wDkU4OYR3utuTp2SZocZazL6XNmn9CM98j7JHKP2Bo4xxQLHOP+lXQivjmKcUvNx2bjEpO7c04jxS2lmHDP+3M7-mzg8YSLD-GgFlAA5JDVqC2Fu8cF51ERfURH13FwnbKbcU4dzjN3Qb8v6cbg02xiMhrR6imm7N+bpAluxY2+t1bQscMQEOxJ6nEs2th465-Cc1Pt4Xad9diMH7mep98mfTY7BRDHNAuK9G4ra3U8g9Tj8Y4wf-Yh7-KHyfX0RiZ279hKv1eEgNDrvXcADdG6nhFU3elzdr1D+HyvJOWtD-axECho+dvV7obTuvDSdbBb0fQw8ABHar1AZAADUeDlHLURiKX3XzUv5MT2fJl49V5FgvMc1+x99boXL4bzvAvjZCTZsOEAOA9Z6LFxBfhatfhWdaPcPE7G-Z6MA2PIoCAx-VRCcOnFPJpdPVpIUbCI4eQXEEQGLI-TqUrdse6InMvB-S-XAEHcMUgs7dOWpR3V-FPZDdPHZdpL3dACwHvclfQATPSITL+MvKMSPJaAQyAuMVZOgq7V9H2LZH2YsVZXAb-X-Jrf-PAkiQA4At6C-Q7HkGfQ7GjcMHQ3bV5BfegyQ27HcXsLWP5U8J7F7N7b4Y5E-DzM-YQ+Am9MvCpe9TQwwlFF-CQhpXsAeZnAIvFA8L3BgKAPdOAdoDgqNAATXvBGT0niKOCvTLzgLIJYmOy8LnwFD5EQMX00V7Fdw1iKLxWwjCR4EsF21i0uT6kuWnRLyoIBz0IGwT3Tm1XEIV3ILTzY0zl8mNR1G4GglIHaDKHKBK2MzKy7Aj0gLmmyJhAnDHBcLIO9V8K6N7EbxKPYQzhwjGLgHyyEjzwLz0nz2lQEhLxYncM+kuJEPaPyJMP8KV16L0VyOSlRFwxiP40sUE0sVkBuNcPOMgMyOAkBPgNgwY06Od17GqmkOqgzVzC924FSmsOexkAiVIEP0M2P0pVjWpRmNcL+3gLv08LaIzVrweMKMYN6NQ0THEkMW9FxGOX72EUETsXmNazLwXWAjJ1uKfWML8M0V9BhM3ThLg1LD3x6ThCmBghqLOTqLOQKXZNaMgOWnv3ZNgw-CQNfV9BQLYx1ILT7EPDEA0GDliywCQT0nNNQRkH+LILmOpzvXtN5Lg3uIFLUwZmZzXyfWwjLDBBIEsDy1wyPRUKtClT6ilVkHSMO1L0gMWOAgMOrzEyTwpLU0CLYyFFCwY3CzgH3UPWOVPXPQLI0JLyTBL0+lLOdLXi1IaSC2aXaTEwgTKA+NMFb01ygA71pG7yZMETN1ZOINmNBLIIoOIXVJrxkFtE3wqGxBYFw3IN1JKNaRVyPBMEUmUmWx0nWwMnpiMjgRlX-F3K5L3PhGZDjHjBFG-3SwIB4BJEeHIO5TAQTHEj1BNNHRDOtDtEnTdBHIWhBGAk-BMhZRjG4HVX-N6GWAAF0gA", #nolint
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```

Note that in a later release, we plan to merge `filesbrowser_block()` and `upload_block()` into
a single variable block.


### Reading data from another stack

`result_block()` allows to read data from another stack. From the below code, the first stack contains
a `dataset_block()` from which the n-th first rows are pulled by the `head_block()`. In the second stack,
we can reuse this smaller dataset. If you dynamically add a third stack, you can also select the third stack
output as input of the second stack (it appears in the result block select input).

```r
library(blockr)

serve_workspace(
  stack1 = new_stack(dataset_block, head_block),
  stack2 = new_stack(result_block),
  title = "My workspace"
)
```

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAIhA9jgGADc4AH1qC3F2VCgCOEEIGRlOPPEARlUZCDhqNNKJPgATKC406PFZDTgoJvbRCT9cQOKG8QAmKpq6sb55dgduUn6TIZGZUjtuOCqhAFlwmUyGbNz8oUC4sABfAF0gA", #nolint
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```
