---
title: "1. Data blocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Data blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)
library(bslib)

note_box <- function(..., color) {
  htmltools::div(
    class = sprintf(
      "m-3 p-2 text-%s-emphasis bg-%s-subtle border border-%s-subtle rounded-3",
      color, color, color
    ),
    ...
  )
}
```

## Introduction

Data blocks are the starting point of any __stack__ and are required by any subsequent __transform__ or __plot__ blocks.

### Package based data block

To add a data set, use the function `new_dataset_block()`, passing a package name to the `package` arg that contains any number of datasets:

```r
new_dataset_block <- function(
  selected = character(),
  package = "datasets",
  ...,
  
) {
  # code  
}
```

The `new_dataset_block()` function defaults to using the base R `datasets` package, but you can supply any valid package containing one or multiple data sets. Here, we use the `{blockr.data}` package to select the `"lab"` data set:

```{r, results="asis", echo=FALSE, warning=FALSE, comment = ""}
# shinylive container
card(
  blockr:::create_app_link(
    blockr:::shinylive_links["dataset-block"],
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)

blockr:::print_shinylive_r_code("dataset-block")
```

```{r, eval=TRUE, echo=FALSE}
note_box(
  "In the registry vignette, we see how to register a block in a cleaner way.",
  color = "primary"
)
```

### Reading data from files

In addition to loading data from a package, blockr also contains a block to upload data from a file.
`{blockr}` exposes two extra data blocks which have to be combined with one 
of the four following data __parser__ blocks:

- `new_xpt_block`, reads xpt files (and sas7bdat files).
- `new_rds_block`, reads rds files.
- `new_json_block`, reads json files.
- `new_csv_block`, reads csv files.

#### Upload data

If you want to load data from any location on your computer, `new_upload_block()` is what you need.
Since the `new_upload_block()` temporarily moves data into a custom location, for security reasons, it might not be always possible.

```{r, results="asis", echo=FALSE, warning=FALSE, comment = ""}
# shinylive container
card(
  blockr:::create_app_link(
    blockr:::shinylive_links["upload-block"],
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)

blockr:::print_shinylive_r_code("upload-block")
```

#### Files browser block

`new_filesbrowser_block()` can read data from the __server file system__.
This means that if no data is available on the server, you won't see anything.
The accessible locations can be customized through the `volumes` parameter like so:

```r
custom_filesbrowser_block <- function(...) {
  new_filesbrowser_block(
    file_path = character(),
    volumes = c(vol1 = "<PATH1>", vol2 = "<PATH2>"),
    ...
  )
}
```

In the following demonstration, dummy dataset has been uploaded to the webR file system so you can play with
the block. But, you can also upload your own data thanks to the corresponding
shinylive feature (upload button).

```{r, results="asis", echo=FALSE, warning=FALSE, comment = ""}
# shinylive container
card(
  blockr:::create_app_link(
    blockr:::shinylive_links["filesbrowser-block"],
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)

blockr:::print_shinylive_r_code("filesbrowser-block")
```

Note that in a later release, we plan to merge `filesbrowser_block()` and `upload_block()` into
a single variable block.


### Reading data from another stack

`new_result_block()` allows you to read data from another stack. From the below code, the first stack contains
a `new_dataset_block()` from which we can select some columns with `new_select_block()`. In the second stack,
we can reuse this smaller dataset. If you dynamically add a third stack, you can also select the third stack's
output as input to the second stack (it appears in the result block select input).

```{r, results="asis", echo=FALSE, warning=FALSE, comment = ""}
# shinylive container
card(
  blockr:::create_app_link(
    blockr:::shinylive_links["result-block"],
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)

blockr:::print_shinylive_r_code("result-block")
```
