---
title: "Developer guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developer guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(DiagrammeR)
```

# `{blockr}` structure

Under the hood, `{blockr}` utilises S3 to creates classes and apply the relevant methods
whenever necessary. We describe the general idea below.

## Introduction

`{blockr}` aims at abstracting the logic behind `{dplyr}` functions
under __blocks__. We assume that a __stack__ is a succession
of dependent blocks. We designed 3 kinds of blocks:

- __data__ block: entry level block to determine input data.
- __transform__ block: materialize `{dplyr}` operations on the selected data.
- __output__ (only plot for now) block: consumes transformed data to produce visualizations.

A __stack__ always starts from a data block and may end either by a data transform blocks or by an output block.
Note that at the moment, there is only one output block per stack.

```{r, echo=FALSE}
mermaid("
flowchart TB
    subgraph stack
      input(Data);
      transform(Transform);
      output(Visualize);
      input-->transform--> output;
    end
")
```

```{r, echo=FALSE}
mermaid("
classDiagram
    class stack
    stack : +name
    stack : +class
    stack : +generate_server(stack, ...)
")
```

The stack is able to dynamically add new blocks through `add_block`.

## Blocks

A block is a structure carrying various information:

- The __class__: `{blockr}` leverages S3, so the class is important to apply
the right methods. For instance, data block, tranform and output blocks have different methods (`generate_server.data_block`, ..., respectively).
- __expr__: is the expression which will be evaluated to produce the block __result__.
- __result__: the block output which can be passed to another block.

```{r, echo=FALSE}
mermaid("
classDiagram
    class block
    block : +name
    block : +expr
    block : +result
    block : +layout
    block : +class
    block : +initialize_block(block)
    block : +is_initialized(block)
    block : +generate_code(block)
    block : +evaluate_block(block, ...)
    block : +generate_server(block, ...)
    block : +update_fields(block, ...)
")
```

A block is composed of __fields__ gathered in a list, which are translated into Shiny inputs.
These fields are the necessary elements to reconstruct the block __expression__ and subsequently
produce the result whenever it gets evaluated.

## Fields

### Simple fields

A field is a structure with a __type__ and a __class__.

```{r, echo=FALSE}
mermaid("
classDiagram
    class field
    field : +type
    field : +class
    field : +initialize_field(x, env)
    field : +validate_field(field)
    field : +update_field(field, value, ...)
")
```

The `validate_field` generic ensures each field has revelant values. For instance, for a string field,
which ultimately becomes a shiny text input, the method `validate_field.string_field` is responsible
for checking that the value is a character string and modifies it if not (I BELIEVE WE'RE DOING
TOO MUCH HERE SINCE THE VALIDATION SHOULD HAPPEN ON THE SHINY SIDE).

### More complex fields

With basic fields, we can't handle __dependencies__ between 2 fields. This is typically what happens
in a join block. To join datasets, we need 2 datasets and some columns for the join. Assuming, 
the dataset choice is managed by a field, changing the datasets may update the possible
choices for the `join_by` field. Below shows the definition of 2 simple fields, one for selecting
the join function (static set of choices) and another for selecting the second dataset.

```r
fields <- list(
  join_func = new_select_field(
    paste(type, "join", sep = "_"),
    paste(join_types, "join", sep = "_")
  ),
  y = new_select_field(y[[1]], y),
  by = ...
)
```

How do we handle the `by` field as it depends on `y`?

This is where we introduce `list_field`. `list_field` accepts a function generating 
the choices with a dependency on other existing fields:

```r
by_choices <- function(data, y) {
    # DEPENDENCY on y
    choices <- intersect(
      colnames(data),
      colnames(eval(as.name(y)))
    )

    ...

    # Return nested field
    list(
      val = new_select_field(
        default,
        choices,
        multiple = TRUE
      )
    )
  }
```

`by_choices` can then be injected inside `new_list_field`:

```r
fields <- list(
  join_func = new_select_field(
    paste(type, "join", sep = "_"),
    paste(join_types, "join", sep = "_")
  ),
  y = new_select_field(y[[1]], y),
  by = new_list_field(sub_fields = by_choices) # HERE
)
```

# Adding new fields

Fields are translated into shiny inputs. For instance, the `select_field` surprinsingly
yields a shiny `selectInput`. It is quite easy to add a new field:

1. Under `./R/field.R`, create a new `new_*_field` and `*_field` functions, `*` being the field name.
2. Under `./R/field.R`, add a specific validation method `validate_field.*_field`.
3. Under `./R/ui.R`, add `ui_input.*_field` and `ui_update.*_field` 
to create the corresponding HTML element for Shiny.

# Create a new block

You may find yourself in a situation where the set of existing blocks does not fully meet
your needs. We exposed few helpers so you can quickly get started creating new block. You can call
`create_block` choosing between `transform` or `plot` block. This creates a new block file
in `R/<name>-block.R`.

If your block has specific output, you may need to tweak `./R/server.R`:

1. Add a `server_output.*_block` method, `*` being the new block name.
2. Add a specific `generate_server.*_block`.

If you had to include new fields, refer to the previous section.
