---
title: "6: Internals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6: Internals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(DiagrammeR)
```

# `{blockr}` structure

Under the hood, `{blockr}` utilises S3 to creates classes and apply the relevant methods
whenever necessary. We describe the general idea below.

## Workspace

TBD

## Stack

The stack is a list of blocks.

```{r, echo=FALSE}
mermaid("
classDiagram
    class stack
    stack : +name
    stack : +class
    stack : +generate_server(stack, ...)
")
```

The stack is able to dynamically add new blocks through `add_block()`.

### Modify a stack

To add a block to a stack, you can leverage `add_block()`:

```r
stack <- new_stack(dataset_block) |>
  add_block(select_block) |>
  add_block(filter_block, 1)
```

If you specify the position (last parameter), you can add a block right 
after the given index. For instance, the above command first inserts
a select block after the data block. Then, a filter block is included
right after the data block.

Note that you can't add a block before the data block and 
before the plot block. In a later version, we may add multiple
plots per stack.

### Moving blocks

TBD

## Blocks

A block is a structure carrying various information:

- The __class__: `{blockr}` leverages S3, so the class is important to apply
the right methods. For instance, data block, tranform and output blocks have different methods (`generate_server.data_block`, ..., respectively).
- __expr__: is the expression which will be evaluated to produce the block __result__.
- __result__: the block output which can be passed to another block.

```{r, echo=FALSE}
mermaid("
classDiagram
    class block
    block : +name
    block : +expr
    block : +result
    block : +layout
    block : +class
    block : +initialize_block(block)
    block : +is_initialized(block)
    block : +generate_code(block)
    block : +evaluate_block(block, ...)
    block : +generate_server(block, ...)
    block : +update_fields(block, ...)
")
```

A block is composed of __fields__ gathered in a list, which are translated into Shiny inputs.
These fields are the necessary elements to reconstruct the block __expression__ and subsequently
produce the result whenever it gets evaluated.

## Fields

### Simple fields

A field is a structure with a __type__ and a __class__.

```{r, echo=FALSE}
mermaid("
classDiagram
    class field
    field : +type
    field : +class
    field : +initialize_field(x, env)
    field : +validate_field(field)
    field : +update_field(field, value, ...)
")
```

The `validate_field` generic ensures each field has revelant values. For instance, for a string field,
which ultimately becomes a shiny text input, the method `validate_field.string_field` is responsible
for checking that the value is a character string and modifies it if not.

### More complex fields

TBD

# Validation and evaluation

That's currently how {blockr} validate blocks and fields. Each block is validated, meaning that each field that compose it should have valid values according to the input data. Only when everything is valid, the block expression is evaluated and the result computed and passed to the next block. Any error is catched by R and forwarded to JavaScript so that users are notified about what went wrong and where.

```{r, echo=FALSE}
mermaid("
flowchart TD;
  subgraph stack
    subgraph blk1
      block1[Block 1] --> |check fields| validate_block1[Validation] --> validate_block1_ok[OK]
      validate_block1 --> validate_block1_fail[Fail]
      validate_block1_ok --> |evaluate| try_eval[Try] --> eval_block1_ok[OK] --> data_block1[Data]
      eval_block1_ok[Success]
      try_eval --> eval_block1_fail[Fail]
    end
    data_block1 -.-> block2
    subgraph blk2
      block2[Block2] --> |check fields| validate_block2[Validation] --> validate_block2_ok[OK]
    
      validate_block2 --> validate_block2_fail[Fail]
      validate_block2_ok --> |evaluate| try_eval2[Try] --> eval_block2_ok[OK] --> data_block2[Data]
      eval_block2_ok[Success]
      try_eval2 --> eval_block2_fail[Fail]
    end
    data_block2 -.-> stack_result[Stack result]
  end
  eval_block1_fail -.-x |notification| gui
  eval_block2_fail -.-x |notification| gui
  validate_block1_fail -.-x |notification| gui
  validate_block2_fail -.-x |notification| gui
  subgraph ui
      gui[User interface]
  end
", height = "1250px") |>
  htmlwidgets::onRender("
    function(el, x) {
      el.classList.add('text-center')
    }
  ")
```

# Adding new fields

Fields are translated into shiny inputs. For instance, the `select_field` surprinsingly
yields a shiny `selectInput`. It is quite easy to add a new field:

1. Under `./R/field.R`, create a new `new_*_field` and `*_field` functions, `*` being the field name.
2. Under `./R/field.R`, add a specific validation method `validate_field.*_field`.
3. Under `./R/ui.R`, add `ui_input.*_field` and `ui_update.*_field` 
to create the corresponding HTML element for Shiny.

# Create a new block

You may find yourself in a situation where the set of existing blocks does not fully meet
your needs. We exposed few helpers so you can quickly get started creating new block. You can call
`create_block` choosing between `transform` or `plot` block. This creates a new block file
in `R/<name>-block.R`.

If your block has specific output, you may need to tweak `./R/server.R`:

1. Add a `server_output.*_block` method, `*` being the new block name.
2. Add a specific `generate_server.*_block`.

If you had to include new fields, refer to the previous section.
