---
title: "new-field"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{new-field}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(blockr)
library(bslib)
```


In this vignette, we show how to create a new field, that is the `slider_field`, similar to the range field with only 1 value.

## Introduction

To create a field, we first need to design a __constructor__. In our case, we call it `new_slider_field`.
Since fields are mirroring shiny inputs, we can naturally provide the corresponding input parameters that are the value, minimum, maximum and step. We pass all of these to `blockr::new_field()` and give it a custom class `slider_field`. 

```{r slider_field}
new_slider_field <- function(
    value = numeric(),
    min = numeric(),
    max = numeric(),
    step = numeric(),
    ...) {
  blockr::new_field(
    value = value,
    min = min,
    max = max,
    step = step,
    ...,
    class = "slider_field"
  )
}
```

On the UI side, this is fairly straightforward as all we have to do is call `shiny::sliderInput` with the same parameters given in the above constructor.

```{r, ui-slider-field}
ui_input.slider_field <- function(x, id, name) {
  ns <- NS(input_ids(x, id))

  shiny::sliderInput(
    blockr::input_ids(x, id),
    name,
    value = blockr::value(x, "value"),
    min = blockr::value(x, "min"),
    max = blockr::value(x, "max"),
    step = blockr::value(x, "step")
  )
}
```

The next step consists in validating the field to ensure provided values are correct. For instance,
we want to make sure all passed elements are numbers. This is done within the `validate_field.slider_field` method. We extract all the field element values with `blockr::value(x, "element")` and use the suitable validate function, that is `validate_number()`. Note that if nothing is provided, the validation falls back to `validate_field.field()`, which does nothing.

```{r, validate-slider-field}
validate_field.slider_field <- function(x) {
  val <- value(x)
  min <- value(x, "min")
  max <- value(x, "max")
  step <- value(x, "step")

  validate_number(val)

  if (length(min)) {
    validate_number(min, "min")
  }

  if (length(max)) {
    validate_number(max, "max")
  }

  if (length(step)) {
    validate_number(step, "step")
  }

  NextMethod()
}
```

Finally, we want to make sure that we can update the field with the `ui_update.slider_field()` method. This is needed whenever another field would have to update the slider field, because of a possible __dependency__ between these 2 fields. As an example, this is what currently happens in the __filter block__ where the selected column field updates the value field (the field type changes depending on the column type).

```{r, update-slider-field}
ui_update.slider_field <- function(x, session, id, name) {
  updateSliderInput(
    session,
    blockr::input_ids(x, id),
    blockr::get_field_name(x, name),
    blockr::value(x),
    blockr::value(x, "min"),
    blockr::value(x, "max")
  )
}
```

## Example

We can test our newly created field in a custom block:

```{r, test-slider-field, eval=FALSE}
new_rows_select_block <- function(...) {
  n_rows <- \(data) nrows(data)

  fields <- list(
    nrows = new_slider_field(
      value = 1,
      min = 0,
      max = n_rows,
      title = "Select rows"
    )
  )

  new_block(
    fields = fields,
    ...,
    expr = quote(data[seq_along(.(nrows)), ]),
    name = "Slider select block",
    class = c("rows_select_block", "transform_block")
  )
}

serve_stack(
  new_stack(
    new_dataset_block,
    new_rows_select_block
  )
)
```
