---
title: "3. Case study: create a plot block"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Case study: create a plot block}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)

note_box <- function(..., color) {
  htmltools::div(
    class = sprintf(
      "m-3 p-2 text-%s-emphasis bg-%s-subtle border border-%s-subtle rounded-3",
      color, color, color
    ),
    ...
  )
}
```

In this vignette, we show how to create a plot block __layer__ by layer with the `{ggplot2}` package.

```{r, eval=TRUE, echo=FALSE}
note_box(
  "Before going further, please be sure to read the S3 chapter of advanced R: https://adv-r.hadley.nz/s3.html.", #nolint
  color = "warning"
)
```

## Introduction

### Anatomy of a `{ggplot2}` plot

With `{ggplot2}`, plots are built __layer__ by __layer__. It all starts with
`ggplot()` which initialises the ggplot object passing optional data and mappings.
Then, we add __geoms__, like `geom_point()`. Geoms also accept custom __mappings__ (overwrite mappings
passed to the first `ggplot()` call), data, as well as other parameters.
As a result, there are many ways to build a ggplot:

```r
# 1
ggplot(data = <DATA>, mapping = <MAPPING>) +
  geom_point()
# 2
ggplot(data = <DATA>) +
  geom_point(mapping = <MAPPING>) +
  geom_point(mapping = <MAPPING>)

# 3
ggplot() +
  geom_point(data = <DATA>, mapping = <MAPPING>) +
  geom_point(data = <DATA>, mapping = <MAPPING>)
```

In this tutorial, we go for the first option.

### To do list

What do we need to create?

It seems obvious to add new `ggplot_block()` and geom __constructors__. On the Shiny side, we have to handle the 
plot output element. Remember that the `server_output()` __generic__ is defined in the `{blockr}`
`server.R` script. It supports tables with `server_output.block()` and plots with `server_output.plot_block()`.
Therefore, we don't need to create another S3 method and have to make sure our new plot block inherits
from the `plot_block` class to __dispatch__ to the correct method. Most of our work will be on the block side since
we'll have to generate the plot layer by layer, each layer being a block. Before writing this vignette,
`{blockr}` did not support layer by layer plots. This new concept required to expose a new way of combining plot expressions
to generate the entire stack code. With the previous infrastructure, we could only do:

```r
data %>% block_1 %>% block_2
```

However, handling `{ggplot2}` grammar requires to be able to combine with `+` such as:

```r
data %>% plot_block + layer_block
```

All the process is described in the following, which should 
highlight the __flexibility__ provided by `{blockr}`.

### New block helpers

```{r, eval=TRUE, echo=FALSE}
note_box(
  "What is described in the following may not be necessary in all cases and 
  is specific to `{ggplot2}` which brings specific syntax like `+`. 
  `{blockr}` now supports layer by layer plots and you theoretically only have 
  to create the corresponding blocks (See next section). 
  What we show below is sort of 'making off' that will serve to advanced users.",
  color = "warning"
)
```

#### New `evaluate_block()` method

We define below a new way to pass data to a block layer, through the `evaluate_block.plot_layer_block()` method.
As a consequence, these block layer must have the `plot_layer_block` class:

```r
evaluate_block.plot_layer_block <- function(x, data, ...) {
  stopifnot(...length() == 0L)
  eval(
    substitute(data + expr, list(expr = generate_code(x))),
    list(data = data)
  )
}
```

Compared to `evaluate_block.transform_block()`, we simply replaced `%>%` by `+`. Nothing more! 
As a side note, if you had to develop your own method in a test script, 
you could register it and test it with:

```r
.S3method("<GENERIC>", "<CLASS>", <METHOD>)
```

#### Generate a valid stack code

Before adding the layer by layer plot blocks, `generate_code.stack()` was:

```r
generate_code.stack <- function(x) {

  binary_substitute <- function(x, y) {
    substitute(x %>% y, list(x = x, y = y))
  }

  Reduce(binary_substitute, lapply(x, generate_code))
}
```

This method leverages `Reduce()` which applies a __binary function__ to combine 
elements of a vector (from left to right by default).
For instance, we could sum the 3 first integers with:

```r
Reduce(`+`, 1:3)
```

`generate_code.stack()` successively combines all the block expressions given by `lapply(x, generate_code)`. 
If there are 3 blocks in the stack:

```r
my_stack <- new_stack(data_block, filter_block, select_block)
```

`binary_substitute` is called twice:

- Once to combine x being `<data_block_EXPR>` (the data block expression) and y being `<filter_block_EXPR>`, 
to give `<data_block_EXPR> %>% <filter_block_EXPR>`.
- Then to combine the previous result, namely x, equal to `<data_block_EXPR> %>% <filter_block_EXPR>` with y equal to `<select_block_EXPR>`.
- This ultimately yields: `<data_block_EXPR> %>% <filter_block_EXPR> %>% <select_block_EXPR>`

`substitute()` ensures we don't evaluate the generated expression and replace `x` and `y` by their respective values.

Can you spot the current point? How do we combine plot layer expressions linked by `+` with this setup?
In order to overcome this limitation, we introduced the `block_combiner()`, 
a new generic which aims at linking multiple block expressions depending on their class. Under the hood,
we only check the class of the __right__ block to determine the expression link (the S3 dispatch occurs on the second block), 
that is `%>%` or `+` (and maybe more in the future). 

```r
block_combiner <- function(left, right, ...) UseMethod("block_combiner", right)
```

Another important modification to bring is about `generate_code.stack()`. 
With the above logic, `Reduce()` only consumes the block expressions, thereby
preventing us from being able to dispatch. We pass a vector of blocks instead, 
such that we get:

```r
generate_code.stack <- function(x) {
  if (length(x) == 0) return(quote(identity()))

  # Handles monoblock stacks
  if (length(x) > 1) {
    aggregate_code <- function(x, y) {
      block_combiner(x, y)
    }
    Reduce(aggregate_code, lapply(x, \(b) b))
  } else {
    generate_code(x[[1]])
  }
}
```

For empty stacks, the returned code is arbitrary. For stacks with one block, we don't 
need `Reduce()` and only call `generate_code()` for the corresponding block. We create a new
internal function, that is `aggregate_code`, subsequently calling `block_combiner()`. 

There is still one last issue to solve. Since we now pass blocks,
we have to find a way to let `block_combiner()` evaluate each block expression. 
We apply `generate_code()` within the `substitute()` call so we can inject the block expression
and not the block itself. This eventually leads us to:

```r
block_combiner.transform_block <- function(left, right, ...) {
  substitute(
    left %>% right,
    list(left = generate_code(left), right = generate_code(right))
  )
}

block_combiner.plot_block <- block_combiner.transform_block

block_combiner.plot_layer_block <- function(left, right, ...) {
  substitute(
    left + right,
    list(left = generate_code(left), right = generate_code(right))
  )
}
```

Importantly, plot constructor like `ggplot()` should have the `plot_block` class
so that the link is `%>%` (the same as for other transform blocks).

### Create the ggplot block

To create a new block, we call the `new_block()` constructor. It expects:

- __fields__: a list of field, which are translated into shiny inputs.
- __expr__: the expression returned by the block, necessary to produce an output and export the code.
- __name__: a name (randomly choosen, you don't need to worry about this).
- __class__: a class to dispatch to the relevant S3 methods.
- __layout__: an optional layout (default to `default_layout_fields()`).

Our block will have `c("ggplot_block", "plot_block")` as classes, some fields to pass mappings and we
leave the layout to the default choice. Note that the mapping field depends on the provided data. 
To stay simple, we assume to only handle `x` and `y` __aesthetics__.

```{r}
new_ggplot_block <- function(data, ...) {
  new_block(
    fields = list(
      # TO DO
    ),
    expr = quote(),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}

ggplot_block <- function(data, ...) {
  initialize_block(new_ggplot_block(data, ...), data)
}
```

We finally end with `new_ggplot_block`, the internal constructor and 
`ggplot_block`, the user facing constructor. The next step is to create the mappings fields with 
two `new_select_field()`. Those fields are converted into shiny inputs, specifically `selectInput()`.
To set the field choices which depend on the data, we define the `data_cols` helper:

```r
data_cols <- function(data) colnames(data)
```

We modify `new_ggplot_block` such that:

```r
new_ggplot_block <- function(data, ...) {

  data_cols <- function(data) colnames(data)

  new_block(
    fields = list(
      x = new_select_field(colnames(data)[1], data_cols, type = "name"),
      y = new_select_field(colnames(data)[2], data_cols, type = "name")
    ),
    expr = quote(),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}
```

Passing `type = "name"` allows to inject functions inside the select field choices. This
is useful to dynamically update the choices whenever data change.

Producing the expression, which is certainly the most "technical" part, as it involves a bit of __metaprogramming__. 
The expression must not be evaluated in the block, that's why it is wrapped in a `quote()`. 
`{blockr}` is then able to generate the expression with `generate_code()` and evaluate it with `evaluate_block()`.

Our ggplot expression can be written as:

```r
ggplot(mapping = aes(x = .(x), y = .(y)))
```

In the above expression, you may notice `.()`, which is actually required by `generate_code()`. Under the hoods, this is needed by `bquote()` which only evaluates arguments wrapped
by `.()` with variables from the environment. It makes it easier to use than `substitute()`, since we explicitly mark what we want to
evaluate.

All combined together, this eventually yields:

```{r, eval=TRUE}
new_ggplot_block <- function(data, ...) {

  data_cols <- function(data) colnames(data)

  new_block(
    fields = list(
      x = new_select_field(colnames(data)[1], data_cols, type = "name"),
      y = new_select_field(colnames(data)[2], data_cols, type = "name")
    ),
    expr = quote(
      ggplot(mapping = aes(x = .(x), y = .(y)))
    ),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}
```

As you can, see the code base is reasonable in terms of complexity.

### Create a geom block

Now that we have a valid `ggplot_block`, we want to add it a geometry. The easiest one is `geom_point()`.
To keep the vignette as simple as possible, we only handle one option, namely the point color. We
host this information in a `new_select_field()`, for instance:

```r
new_select_field("blue", c("blue", "green", "red"))
```

The expression is straightforward to get:

```r
geom_point(color = .(color)) # Don't forget to wrap it with quote(...).
```

Importantly, since we use `.(color)`, the field name must be __color__.

This finally gives us:

```{r, eval=TRUE}
new_geompoint_block <- function(data, ...) {

  new_block(
    fields = list(
      color = new_select_field("blue", c("blue", "green", "red"))
    ),
    expr = quote(
      geom_point(color = .(color))
    ),
    class = c("plot_layer_block", "plot_block"),
    ...
  )
}

geompoint_block <- function(data, ...) {
  initialize_block(new_geompoint_block(data, ...), data)
}
```

Note the class `plot_layer_block`. This is necessary to invoke the corresponding 
`evaluate_block()` method (to use `+` instead of `%>%`).


### Try it

```{r, eval=TRUE, echo=FALSE}
note_box(
  "We currently decided that each plot layer can be visible, so that you can
  see the plot construction step by step within the UI. In any case, `{blockr}` 
  only expands the latest stack block UI so you can focus on the final result.
  Toggling other blocks is up to the end-user.",
  color = "primary"
)
```

We can try it on the following stack:

```r
stack <- new_stack(data_block, ggplot_block, geompoint_block)
serve_stack(stack)
```

### Going further

Below is a possible implementation of an interactive `{ggiraph}` [plot](https://davidgohel.github.io/ggiraph/#with-r-and-r-markdown).
<p class="text-center">
<a 
class="btn btn-primary" 
data-bs-toggle="collapse" 
href="#ggiraphBlockr" 
role="button" 
aria-expanded="false" 
aria-controls="ggiraphBlockr">
Toggle code
</a>
</p>

<div class="collapse" id="ggiraphBlockr">

```r
library(blockr)
library(ggplot2)
library(ggiraph)

custom_data_block <- function(...) {
  initialize_block(
    new_data_block(
      ...,
      selected = "mtcars"
    )
  )
}

new_ggplot_block <- function(data, ...) {

  data_cols <- function(data) colnames(data)

  new_block(
    fields = list(
      x = new_select_field("wt", data_cols, type = "name"),
      y = new_select_field("qsec", data_cols, type = "name"),
      color = new_select_field("disp", data_cols, type = "name")
    ),
    expr = quote(
      ggplot(mapping = aes(x = .(x), y = .(y), color = .(color)))
    ),
    class = c("ggplot_block", "plot_block"),
    ...
  )
}

ggplot_block <- function(data, ...) {
  initialize_block(new_ggplot_block(data, ...), data)
}

# We could have use a mutate_block instead of
# changing the data from inside the block ...
# {blockr} is flexible here.
new_geompoint_interactive_block <- function(data, ...) {
  build_expr <- function(data) {
    # Get data from the previous ggplot layer
    # data is the ggplot so we need to use data$data
    # to get the initial data
    data$data$carname <- row.names(data$data)
    substitute(
      geom_point_interactive(
        # Pass in new data
        data = new_data,
        aes(
          tooltip = carname,
          data_id = carname
        )
      ),
      list(new_data = data$data)
    )
  }

  new_block(
    fields = list(
      expression = new_hidden_field(build_expr)
    ),
    expr = quote(.(expression)),
    class = c("plot_layer_block", "plot_block"),
    ...
  )
}

geompoint_interactive_block <- function(data, ...) {
  initialize_block(new_geompoint_interactive_block(data, ...), data)
}

new_theme_block <- function(data, ...) {
  new_block(
    fields = list(
      theme = new_select_field(
        "theme_minimal", 
        grep("^theme_.*$", ls("package:ggplot2"), perl = TRUE, value = TRUE),
        type = "name"
      )
    ),
    expr = quote(
      .(theme)()
    ),
    class = c("plot_layer_block", "plot_block"),
    ...
  )
}

theme_block <- function(data, ...) {
  initialize_block(new_theme_block(data, ...), data)
}

new_ggiraph_block <- function(data, ...) {
  new_block(
    fields = list(
      pointsize = new_numeric_field(12, min = 1, max = 20)
    ),
    expr = quote(
      girafe(ggobj = data, pointsize = .(pointsize))
    ),
    class = c("ggiraph_block", "plot_block"),
    ...
  )
}

ggiraph_block <- function(data, ...) {
  initialize_block(new_ggiraph_block(data, ...), data)
}

stack <- new_stack(
  custom_data_block,
  ggplot_block,
  geompoint_interactive_block,
  theme_block,
  ggiraph_block
)
serve_stack(stack)
```

</div>
