---
title: "2. Blocks registry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Blocks registry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)
library(bslib)
library(DiagrammeR)

note_box <- function(..., color) {
  htmltools::div(
    class = sprintf(
      "m-3 p-2 text-%s-emphasis bg-%s-subtle border border-%s-subtle rounded-3",
      color, color, color
    ),
    ...
  )
}
```

## Introduction

The __registry__ is a __environment__ which provides access to multiple __blocks__ as well as some
metadata:

- The block __description__.
- Allowed __input__ (default to data.frame, expect for `new_dataset_block` blocks)
- Returned __output__.
- ...

In other words, the __registry__ is a __"supermarket"__ for data analysis. As shown below, if you develop your own blocks package and registers blocks on load, these blocks become available to the end user. Therefore this makes it powerful for __collaboration__ between data science teams.

```{r, echo=FALSE}

mermaid("
  flowchart LR
  subgraph blockr_ggplot2[blockr.ggplot2]
    new_block1[New block]
    new_block2[New block]
  end
  subgraph blockr_echarts4r[blockr.echarts4r]
    new_block3[New block]
    new_block4[New block]
  end
  blockr_ggplot2 --> |register| registry
  blockr_echarts4r --> |register| registry
  subgraph registry[Registry]
    subgraph select_reg[Select block]
      reg_name[Name: select block]
      reg_descr[Description: select columns in a table]
      reg_classes[Classes: select_block, tranform_block]
      reg_input[Input: data.frame]
      reg_output[Output: data.frame]
      reg_package[Package: blockr]
    end
    subgraph filter_reg[Filter block]
    end
    filter_reg --x |unregister| trash['fa:fa-trash']
  end
  ",
  height = "600px"
) |>
  htmlwidgets::onRender(
    "function(el, x) {
      el.classList.add('text-center')
    }
    "
  )
```

## Previewing available blocks

Upon loading, `{blockr}` __registers__ its internal __blocks__ with `register_blockr_blocks()`.
You won't have to call this function as it is not exported anyway. This makes the __registry__ environment
ready to be queried by `available_blocks()`. A truncated output example below:

```r
$dataset_block
function(...) {
  ...
}
<environment: namespace:blockr>
attr(,"name")
[1] "data block"
attr(,"description")
[1] "Choose a dataset from a package"
attr(,"classes")
[1] "dataset_block" "data_block"   
attr(,"input")
[1] NA
attr(,"output")
[1] "data.frame"
attr(,"package")
[1] "blockr"
attr(,"class")
[1] "block_descr"

$select_block
function(data, ...) {
  ...
}
<environment: namespace:blockr>
attr(,"name")
[1] "select block"
attr(,"description")
[1] "select columns in a table"
attr(,"classes")
[1] "select_block"    "transform_block"
attr(,"input")
[1] "data.frame"
attr(,"output")
[1] "data.frame"
attr(,"package")
[1] "blockr"
attr(,"class")
[1] "block_descr"
```

```{r, eval=TRUE}
names(available_blocks())
```

## Register a block
To register your own blocks, user facing functions are:

- `register_block()` to register a block in the __registry__. If the __block__ is already registered,
it __overwrites__ the existing one.
- `register_blocks()` to register multiple blocks.

Let's say you want to create a new `new_tail_block` which returns the `n` last rows of the
selected dataset:

```{r, eval=TRUE}
new_tail_block <- function(data, n_rows = numeric(), ...) {
  n_rows_max <- function(data) nrow(data)

  new_block(
    fields = list(
      n_rows = new_numeric_field(n_rows, 1L, n_rows_max)
    ),
    expr = quote(tail(n = .(n_rows))),
    class = c("tail_block", "transform_block"),
    ...
  )
}

register_block(
  constructor = new_tail_block,
  name = "tail block",
  description = "return last n rows",
  category = "transform",
  classes = c("tail_block", "transform_block"),
  input = "data.frame",
  output = "data.frame"
)
```

If we now query the registry, the new block is available:

```{r, eval=TRUE}
names(available_blocks())
```

We can now run the demo app to add blocks. The new `tail_block` should be
visible in the new choices:

```{r, eval=TRUE, echo=FALSE}
card(
  blockr:::create_app_link(
    blockr:::shinylive_links["registry-demo"],
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```

If you had to register multiple blocks:

```r
register_blocks(
  constructor = c(
    new_dataset_block, new_filter_block, new_select_block, new_summarize_block
  ),
  name = c(
    "data block", "filter block", "select block", "summarize block"
  ),
  description = c(
    "choose a dataset in a package",
    "filter rows in a table",
    "select columns in a table",
    "summarize data groups"
  ),
  category = c("data", rep("transform", 3)),
  classes = list(
    c("dataset_block", "data_block"),
    c("filter_block", "transform_block", "submit_block"),
    c("select_block", "transform_block"),
    c("summarize_block", "transform_block", "submit_block")
  ),
  input = c(NA_character_, "data.frame", "data.frame", "data.frame"),
  output = c("data.frame", "data.frame", "data.frame", "data.frame"),
  package = "<PKG_NAME>"
)
```

## Unregister a block

The counterpart of `register_block()` is `unregister_blocks()`. We can remove
our new `tail_block` from the registry:

```{r, eval=TRUE}
unregister_blocks(ids = "tail_block")

# Check it out
names(available_blocks())
```

where __ids__ is the first entry we applied in the class attributes when the block
was created, that is `tail_block`. If you had to remove multiple blocks at once,
you can pass a vector in __ids__.
