---
title: "2. Blocks registry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Blocks registry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(blockr)
library(bslib)
```

## Introduction

This part is likely more dedicated to developers than users. However,
it might be useful whenever you design custom blocks and want to register
them in `{blockr}`.

The registry is a __environment__ which provides access to multiple blocks as well as some
metadata:

- The block description.
- Inputs accepted (default to data.frame, expect for `data_block` blocks)
- Output returned.
- ...

## Previewing available blocks

Upon loading, `{blockr}` registers its internal blocks with `register_blockr_blocks()`.
You won't have to call this function as it is not exported anyway. This makes the registry environment
ready to be queried by `available_blocks()`. A truncated output example below:

```r
$dataset_block
function(...) {
  initialize_block(new_data_block(...))
}
<environment: namespace:blockr>
attr(,"name")
[1] "data block"
attr(,"description")
[1] "choose a dataset"
attr(,"classes")
[1] "dataset_block" "data_block"   
attr(,"input")
[1] NA
attr(,"output")
[1] "data.frame"
attr(,"package")
[1] "blockr"
attr(,"class")
[1] "block_descr"

$select_block
function(data, ...) {
  initialize_block(new_select_block(data, ...), data)
}
<environment: namespace:blockr>
attr(,"name")
[1] "select block"
attr(,"description")
[1] "select columns in a table"
attr(,"classes")
[1] "select_block"    "transform_block"
attr(,"input")
[1] "data.frame"
attr(,"output")
[1] "data.frame"
attr(,"package")
[1] "blockr"
attr(,"class")
[1] "block_descr"
```

```{r, eval=TRUE}
names(available_blocks())
```

## Register a block
To register you own blocks, user facing functions are:

- `register_block()` to register a block in the registry. If the block is already registered,
it overwrites the existing one.
- `register_blocks()` to register multiple blocks.

Let's say you want to create a new `head_block` which returns the `n` first rows of the
selected dataset:

```{r, eval=TRUE}
new_head_block <- function(data, n_rows = numeric(), ...) {

  new_block(
    fields = list(
      n_rows = new_numeric_field(n_rows, 1L, 100L)
    ),
    expr = quote(head(n = .(n_rows))),
    class = c("head_block", "transform_block"),
    ...
  )
}

register_block(
  constructor = new_head_block,
  name = "head block",
  description = "return first n rows",
  classes = c("head_block", "transform_block"),
  input = "data.frame",
  output = "data.frame"
)
```

If we now query the registry, the new block is available:

```{r, eval=TRUE}
names(available_blocks())
```

We can now run the demo app to add blocks:

```{r, eval=TRUE, echo=FALSE}
card(
  create_app_link(
    "NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YOtyIEA1gyG4ABAzioi7GQF4ZQgBalSqdogD0+kWMkBaIgwDmGSy1IaArnQwsi+2o1PzF7IQEoBCG4WRigGAE8+YwkGAKCQhjDI9g02cLjg0Iio0RiMABMoLjjAiDhqAH0NOCh8iujxGQAeUxkAMwcIAlJXCD5CrlkICoYiamU1CAd4BhYCPj9ZDGW-GRBAmRkyyobBCE3NtpY4bnyJmWDOPYOD4dHx1S3yiqmZuYqjk-y+O7H2WQAjAAZQEABlBQLiN0WGwOcAAHqgGI8AI4OIjkPjVWo-R4YH4jP5+Ym4WGbAjcKDsc7zTQ1OoNaTqMCkRIcNoWGD1XLifyk-YHZYYWFxAC+gUC8lsnDgDG5JmuMmIHFZDm6Fke2yq9PlEn5m2g8EedNqMkZeFh+Tg7AIs1QPRIxrA8lIDgY+yODE4Wzkf2ksIpVPY1setLA2IZPKZQlZUHZnN1vLAMIFbFQDlIToGUAwbUS8H9AqIGfTmbUQmzufzcCEgRKKqgEmarS1nEb4n6RSgibiKTSAEF0IqHCxHnR2JlmKgoJY4B9uCPvmSZLVIyYKiO+CwrWQ7Ol9QdZ2VEuQNyw+G29TJt06YOFL0moTIU5tgwwAG6yx4dLoOvppjNZGLHRAJkYNqV6VZ1gFTZ3x4ZQWjkGpuhYT8ADUeAca0fmeBpHgAOQAVSBSFlwfZsZCPWUijnN9PwYRUbgfA8bhvcswDvB9CxuA1cJ5R55EbHpPz4ODuHYAASLUGifTZ6xuIhx1lT8AFFPzILcIFLCTV1kaCeJkMTJOk-jEKIkjl02ABiGR+24agoHvFd0AofIZFIIgwK4CRLIuMQKO4ChLHsC9vPECSGnYWSDjwxCoDglhKREOdIoWYBgAA0gJODQLujgNcJAAXUK3ybNXRMZARVA4G6dzqiVN15DILz21kew4DNfiPJXfI3LjNy7GUHw7F6UqZD4cygWvZQjjKdoNXamRhr-WQ2DqoolSpDrFrw6hEu4MbLE8jbFtc6LYPgqS+JMAKWCuXzNjwtQGhYnjlt6R5cl86LRSfcUID8MBRUKoA", #nolint
    "app",
    header = FALSE
  ),
  full_screen = TRUE
)
```

If you had to register multiple blocks:

```r
register_blocks(
  constructor = c(
    data_block, filter_block, select_block, summarize_block
  ),
  name = c(
    "data block", "filter block", "select block", "summarize block"
  ),
  description = c(
    "choose a dataset",
    "filter rows in a table",
    "select columns in a table",
    "summarize data groups"
  ),
  classes = list(
    c("dataset_block", "data_block"),
    c("filter_block", "transform_block", "submit_block"),
    c("select_block", "transform_block"),
    c("summarize_block", "transform_block", "submit_block")
  ),
  input = c(NA_character_, "data.frame", "data.frame", "data.frame"),
  output = c("data.frame", "data.frame", "data.frame", "data.frame"),
  package = pkg
)
```

## Unregister a block

The counterpart of `register_block()` is `unregister_blocks()`. We can remove
our new `head_block` from the registry:

```{r, eval=TRUE}
unregister_blocks(ids = "head_block")

# Check it out
names(available_blocks())
```

__ids__ is the first entry we applied in the class attributes when the block
was created, that is `head_block`.